Introduction to *cellcount*

### 2022-09-27 cellcount_intro.Rmd
Updates

```{r, message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!require("cellcount")) {
  BiocManager::install("cellcount")
}
library(cellcount)
```

@@

```{r}
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
```

@@

```{r}
grey_images <- lapply(img_transposed, greyscale, contrast = 1)
display(grey_images[[5]])
```

@@

```{r}
imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
```

@@

```{r}
imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

@@

```{r}
Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
Cell.Count$SD_range<-outlier()
Cell.Count
```

@@

The following code uses these variables to calculate cell density and adds 
the data into the "Cell.Count" data frame:

```{r}
cell.den <- cell_density(cell.total, height = 203.2, width = 166.4, diameter = 2.5, images = 15,
                         filtration.area = 491, volume = 1, total.volume = 100)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
```

@@

```{r}
write.csv(Cell.Count, paste0(savdir, "/Alexandrium counts.csv"))
```

>>>>

if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!require("cellcount")) {
  BiocManager::install("cellcount")
}
library(cellcount)

@@

```{r}
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/Microcystis_F192_40x/", pattern = "tif", full.name = F)
images <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = T)
images_names <- list.files("C:/Users/Tyler.Harman/Desktop/cellcount_work/cellcount_main/test_images/Microcystis_F192_40x/", pattern = "tif", full.name = F)
```

@@

```{r}
greyscale <- function(x, contrast = 2) {
  x <- contrast * x
  x <- x[, , 1] + x[, , 2] + x[, , 3]
  x <- x / max(x)
  x <- normalize(x, inputRange = c(0.1, 0.75))
  return(x)
}
grey_images <- lapply(img_transposed, greyscale, contrast = 1)
display(grey_images[[5]])
```

@@

```{r}
mapped <- function(x, threshold = 0.3) {
  x <- as.matrix(x)
  x[x < threshold] <- 0
  return(x)
}
imagesMapped <- lapply(grey_images, mapped, threshold = 0.3)
```

@@

```{r}
border <- function(x, y) {
  c(x[1:y[1], 1], x[1:y[1], y[2]], x[1, 1:y[2]], x[y[1], 1:y[2]])
}
countCells <- function(x) {
  countAll <- max(x)
  dims <- dim(x)
  border1 <- border(x, dims)
  ids <- unique(border1[which(border1 != 0)])
  countInner <- countAll - length(ids)
  return(countInner)
}
countImages <- function(image, normalize = TRUE, removeEdgeCells = TRUE) {
  if (removeEdgeCells) {
    dims <- dim(image)
    border1 <- border(image, dims)
    ids <- unique(border1[which(border1 != 0)])
    inner <- rmObjects(image, ids)
    EBImage::colorLabels(inner, normalize = normalize)
  } else {
    EBImage::colorLabels(image, normalize = normalize)
  }
}
image_convert <- function(x, w = 17, h = 17, offset = 0.001, areathresh = 50, tolerance= 0.5, ext = 1) {
  image <- thresh(x, w = w, h = h, offset = offset)
  image1 <- fillHull(image)
  image2 <- watershed(distmap(image1), tolerance = tolerance, ext = ext)
  nf <- computeFeatures.shape(image2)
  nr <- which(nf[, "s.area"] < areathresh)
  image3 <- rmObjects(image2, nr)
  return(image3)
}
imagesConverted <- image_convert(imagesMapped[[5]], w = 10, h = 10, offset = 0.001, areathresh = 50, tolerance = 0.8, ext = 1)
final_img <- countImages(imagesConverted, normalize = T, removeEdgeCells = T)
display(final_img)
```

@@

```{r}
Cell.Count$Cell_Count<-as.numeric(Cell.Count$Cell_Count)
outlier<-function(){
  mean<-median(Cell.Count$Cell_Count)
  sd<-sd(Cell.Count$Cell_Count)*2
  sd_neg<-(mean-sd)
  sd_pos<-(mean+sd)
  SD_range_detection<-ifelse(sd_pos>Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)&
    ifelse(sd_neg<Cell.Count$Cell_Count,Cell.Count$Cell_Count,NA)
  return(SD_range_detection)
}
Cell.Count$SD_range<-outlier()
Cell.Count
```

@@

Field of view (FOV) is an important variable calculated from image height,
image width, and cell diameter variables. The following equation is listed
below:

$FOV = HW - 2D(H+W)+4D^{2}$

Where:

$H = height$

$W = width$

$D = diameter$

Image height and width are variables that are taken from dimensions 
disseminated by your microscope camera's manufacturer. Please refer to their 
documentation to calculate your FOV variable correctly.

With the variables calculated, the following code uses these variables to 
calculate cell density and adds the data into the "Cell.Count" data frame:

```{r}
cell_density <- function(x, FOV = 0.0726, images = 10, filtration.area = 213.8, volume = 1, total.volume = 5) {
  MCF <- filtration.area / FOV
  N <- x / images
  cyano <- (MCF * N) / volume
  cyano_total <- cyano * total.volume
  return(cyano_total)
}
cell.den <- cell_density(cell.total, FOV = 0.00084, images = 10, filtration.area = 213.8, volume = 0.05, total.volume = 2)
Cell.Count[nrow(Cell.Count) + 1, ] <- c("Cell Density", cell.den,"null")
```

@@

```{r}
write.csv(Cell.Count, paste0(savdir, "/F192_Test counts.csv"))
```

# Potential errors

End-users should note that potential errors with image analysis using
*cellcount* may occur. Here, we outline a few examples of what may occur, and
how to combat these errors to ensure the highest image quality for accurate
enumeration results.

## Extreme detail from fluorescence

While high image quality is applauded when using this package, sometimes the
natural auto-fluorescence may be too detailed, revealing potential problems
when analyzing images. End-users may want to try different objectives in order
to reduce detail as much as possible while able to show distinction between
cells, or use another fluorescent dye (i.e. DAPI) in order to the program to
produce more accurate enumeration results.

![A greyscale image (left) of *Alexandrium catenella* and a binary filter 
image (right) demonstrating the error results of high-detail from cells,
causing incorrect enumeration results.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\Error1.png)

![A color-label image of *Prorocentrum* sp. where high details were shown from
its fluorescing outer membrane, which caused the program to divide and 
enumerate distinct regions shown here.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\3214435.png)

## High background intensities

If high background intensities is observed within images, errors could occur
like the one shown below, where the threshold adjustment built into the
"mapped" function cannot separate the high intensity background from the
majority of cells within the image. End-users should re-take these images and
reduce the background intensity as much as possible. Snap images as rapidly as
you can to avoid fluorescent dyes from increasing in fluorescent output - one
of the contributing factors that we have observed throughout the development
of this package.

![A color-label image of *Shewanella* IRI-160 resulting from a high
background intensity error. This caused the program to produce an 
inconclusive image, due to the lack of being able to separate the high
background intensity from the cells within the image.](C:\Users\Tyler.Harman\Desktop\cellcount_work\misc_\Error Images\542365.png)
